Final paper: Towards Proving Termination of \texttt{zelda-mosaic} Algorithms under Conditions on the Input

### 2021-05-02 paper/paper.tex
paper: copy overleaf diffs over

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{array,booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[inline]{enumitem}
\newlist{inlist}{enumerate*}{1}
\setlist[inlist]{label= (\alph*)}
\usepackage{mathpartir,semantic}
\usepackage[colorlinks]{hyperref}
\hypersetup{%
    colorlinks=true,
    linkcolor=blue,
    urlcolor=cyan,
    citecolor=blue
}
\usepackage[style=ieee]{biblatex}
\addbibresource{sources.bib}

@@

\section{Progress Report}

The starting point to our project was to fill in some knowledge gaps that we had
not covered in class prior to when we began. Since we planned to model a
program, we needed to know more about the modeling of programs and execution. To
this end, we completed work in \textit{Software Foundations} up to the chapter
\textit{Simple Imperative Programs}~\cite{Pierce:SF1}. This chapter provided
great insight on how to model statements and expressions, maintain state over a
program's execution, and define semantics. After completing this, we were better
able to outline the scope of the project.

Our progress thus far can be divided into two main categories:
\begin{inlist}
\item determining the subset of \matlab\ (dubbed \mmatlab) relevant to our program, and
\item creating data structures and functions suitable to model the behaviors of
    \mmatlab\@.
\end{inlist}

We also identify two main future tasks that will rely upon the completion of
these tasks:
\begin{inlist}
\item binding operations from the \mmatlab\ language to semantics in
    Coq~\cite{Coq}, and
\item using Hoare logic~\cite{Hoare_1969} to prove that the program terminates
    under what pre-conditions.
\end{inlist}

@@

We give a brief overview on \mmatlab, the subset of \matlab\ that we model in
order to prove facts about our programs.

Aside from general programming patterns for control flow, the main syntax of
relevance in our program involves array/matrix accesses and updates, arithmetic
operations, and \matlab-specific primitives such as ranges. We have identified
all of the expressions and statements we expect to use and defined them as
inductive types, shown in \figurename~\ref{fig:inductivetypeexp} and
\figurename~\ref{fig:inductivetypestmt}.

@@

    \item \textsf{AddExp}: While this is nominally a regular addition operator,
        \matlab's addition is interesting in that it has overloaded parameter
        types and expected behavior for each. The addition of two numbers is a
        number and the addition of two matrices results in element-wise
        addition, as one might expect. However, the sum of a number and a
        matrix, is a matrix with all elements incremented by the number. Similar
        behaviors hold for \textsf{MultExp} and \textsf{DivExp}.

    \item \textsf{RangeExp}: Like Python, \matlab\ has an expression where a
        starting number and stopping number may be provided, generating an
        inclusive list of all numbers between them.

@@

\subsection{Hoare Logic}

We have a significant amount of work to do in modeling \mmatlab, but the
blue-sky goal remains to prove the termination of our program given that certain
preconditions hold. Ben has been studying Hoare logic~\cite{Hoare_1969}, the
relevant background material for this problem, as part of his Thesis
requirement. What will be possible to prove will be determined in part by what
we are able to successfully model, given the time constraints. However, as a
preliminary step, we have studied partial Hoare logic and total Hoare logic to
provide more context for the task ahead of us. In this exploration, we found
that partial Hoare logic is a weak statement about programs; namely, given
preconditions \(P\) and postconditions \(Q\) and a program \(C\), the Hoare
triple \(\{P\}C\{Q\}\) asserts that \emph{if} \(C\) terminates, then the
pre-conditions imply the post-conditions. Total Hoare logic, by contrast, is a
stronger version of this claim: \(C\) is asserted to terminate. Thus a proof in
total Hoare logic necessitates a proof that \(C\) terminates; the inference
rules are accordingly adjusted. The primary adjustment is to the rule for while
loops, which are guaranteed to terminate by relation to a decreasing sequence
(with respect to a well-founded order). Given that we are trying to prove
termination, it seems likely that we will need to use total Hoare logic in the
relevant proofs.

\begin{figure}[ht]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
        \FOR{\(tile \in tiles\)}
            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
            \ENDFOR
            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
            \STATE{\(best\_index \gets best\_indices_0\)}
            \STATE{\(tile \gets thumbnails_{best\_index}\)}
        \ENDFOR
    \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 1}\label{alg:v1}
\end{figure}

\begin{figure}[ht]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
        \STATE{\(used \gets \emptyset\)}
        \STATE{\(threshold \gets 1\)}
        \IF{\(|thumbnails| < |tiles|\)}
            \STATE{\(threshold \gets \lceil{|tiles| / |thumbnails|}\rceil\)}
        \ENDIF
        \FOR{\(tile \in tiles\)}
            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
            \ENDFOR
            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
            \STATE{\(best\_index \gets best\_indices_0\)}
            \WHILE{\(used_{best\_index} \ge threshold\)}
                \STATE{\(mses_{best\_index} \gets \infty\)}
                \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
                \STATE{\(best\_index \gets best\_indices_0\)}
            \ENDWHILE
            \STATE{increment \(used_{best\_index}\)}
            \STATE{\(tile \gets thumbnails_{best\_index}\)}
        \ENDFOR
        \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 2}\label{alg:v2}
\end{figure}

\begin{figure}[ht]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{}\COMMENT{Repeat Version 2}
        \STATE{\(tiles \gets \textrm{Version2}(key\_art, thumbnails, size)\)}
        \STATE{\(tiles \gets \textrm{imgaussfilt}(tiles)\)}
        \STATE{\(strip\_size \gets \lfloor{size / 8}\rfloor\)}
        \STATE{}\COMMENT{Blur by averaging}
        \FOR{horizontal \(strip \in tiles\)}
            \STATE{\(strip \gets \textrm{mean}({strip})\)}
        \ENDFOR
        \FOR{vertical \(strip \in tiles\)}
            \STATE{\(strip \gets \textrm{mean}({strip})\)}
        \ENDFOR
        \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 3}\label{alg:v3}
\end{figure}

@@

>>>>

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{array,booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usetikzlibrary{matrix,calc}
\newlist{inlist}{enumerate*}{1}
\setlist[inlist]{label= (\alph*)}
\usepackage{mathpartir,semantic}
\usepackage[colorlinks]{hyperref}
\hypersetup{%
    colorlinks=true,
    linkcolor=blue,
    urlcolor=cyan,
    citecolor=blue
}
\usepackage[style=ieee]{biblatex}
\addbibresource{sources.bib}

@@

\section{Problem Description}
Many non-trivial programs make use of arrays, and our \texttt{zelda-mosaic} program is no different. Because our goal was to prove the termination of the program when given a well-formed input and the core data structure modified throughout the program was an image (represented as an array), it is helpful to first describe the array operations used and the invariants of an array that contribute to ``well-formed-ness".

\subsection{Array Usage}
\matlab's support for multi-dimensional arrays makes it amenable to image-processing tasks such as ours. The algorithms we wrote are described at high-level in pseudocode in Figures \figurename~\ref{alg:v1}, \ref{alg:v2}, and \ref{alg:v3}, but here we explain them in layman's terms to help motivate some features of our proof.\\

As described in our pseudocode, the inputs to our program were a piece of \textit{key art} to be recomposed with \textit{thumbnails} of a certain \textit{size}, defined as the width and height of the thumbnails (all have been coerced to be square-shaped). The original representation of \textit{key art} in \matlab is a two-dimensional array (with indices corresponding to a particular pixel's row and column) containing an additional array of three values - one each for the red, green, and blue color channels. So far, this is an array of three dimensions. We then divide this image into \textit{chunks}, where each \textit{chunk} will be replaced by a \textit{thumbnail} in the final product. For \textit{key art} of dimensions $m \times n$ (with $m$ corresponding to and $n$ corresponding to columns), each image will be divided into $(m / size) \times (n / size)$ \textit{chunks}, which will each be replaced by a \textit{thumbnail} as appropriate. These \textit{chunks}, each of size $size \times size$, are then stored in a two-dimensional array, with each \textit{chunk} retaining its original placement in the \textit{key art}. As a result we now have a five-dimensional array, and in order to access a particular value from a color channel, we need the following values:

\begin{itemize}
\item the \textit{chunk row} it is in
\item the \textit{chunk column} it is in
\item the \textit{x-offset} of the pixel into its chunk
\item the \textit{y-offset} of the pixel into its chunk
\item the \textit{color channel} desired out of red, green, and blue
\end{itemize}

Thus, once a piece of \textit{key art} is successfully divided into \textit{chunks}, the image is effectively represented as an array whose dimensions correspond to the above values. The \textit{chunks} are iterated through in row-major order, gradually being replaced in-place by \textit{thumbnails}. Consequently, the state of our core array maintains some matrix-like properties that are supported by \matlab\@ arrays, but not particular to \matlab\@ in general:

\begin{itemize}
    \item The shape of a matrix \textbf{$M$} can be described by a list of \textit{d} dimensions, where each dimension's value $v_i$ refers to the number of elements held along that dimension. For example, an array with dimensions $3 \times 4 \times 2$ contains 3 arrays, each of which contains 4 arrays of 2 elements. $v_1 = 3$, $v_2 = 4$, and $v_3 = 2$.
    \item Once the shape of \textbf{$M$} is defined and all of the elements have been populated, the shape will not change and all of the elements will remain populated. For example, the aforementioned $3 \times 4 \times 2$ array must always contain 3 arrays, each of which contains 4 arrays of 2 elements.
    \item Each $i$-th dimension from $1$ to $d-1$ must contain $v_i$ elements, each of which is an array of length $v_{i+1}$; in other words, the length of each element in a particular dimension must be identical to one another.
\end{itemize}

In summary, these properties enforce a notion of ``well-formed-ness" that ensures that the size $v_i$ of each dimension remains fixed and that all element in a particular dimension $i$ have the same length as one another. This property is generally true of matrices, hence the references to both an ``arrays" and ``matrices" in this section. Unless otherwise noted, ``array" refers to the generic \matlab\@ object and ``matrix" refers to an ``array" with these properties when discussed in the scope of this paper.

\subsection{Operations} \matlab\@ supports a number of operations that can be used to manipulate arrays, but only a subset are used in our program. To illustrate the operations used, we begin with the following visualization of a three-dimensional array \textbf{$E$} with dimensions $3 \times 4 \times 2$, as it is used in \matlab:\\

% Code adapted from: https://tex.stackexchange.com/questions/300109/simple-visualization-of-3d-matrix

\begin{figure}[h!]
    \centering
    \begin{tikzpicture}[every node/.style={anchor=north east,fill=white, minimum width=.7cm}]
    \matrix (mA) [draw,matrix of math nodes]
    {
    1 & 3 & 5 & 7 \\
    2 & 4 & 6 & 8 \\
    };

    \matrix (mB) [draw,matrix of math nodes] at ($(mA.south west)-(-1.5, .1)$)
    {
    9 & 11 & 13 & 15 \\
    10 & 12 & 14 & 16 \\
    };

    \matrix (mC) [draw,matrix of math nodes] at ($(mB.south west)-(-1.5, .1)$)
    {
    17 & 19 & 21 & 23 \\
    18 & 20 & 22 & 24 \\
    };

    \draw[dashed](mA.north east)--(mC.north east);
    \draw[dashed](mA.north west)--(mC.north west);
    \draw[dashed](mA.south east)--(mC.south east);
    \end{tikzpicture}

\end{figure}

\subsubsection{Element Access}
For the matrix \textbf{$E$}, each element would be indexed according to the following mapping. Under this scheme, for example, the element $8$ in \textbf{$E$} would have the index $(1, 4, 2)$. This indexing scheme is the general pattern used to access one particular element out of a multi-dimensional array. Note that arrays in \matlab\@ have indices starting at 1.

\begin{figure}[h!]
    \centering
    \begin{tikzpicture}
    \def\xs{1} %shift in x direction
    \def\ys{1.5} %shift in y direction
    \def\nm{3} % number of 2d matrices in the 3d matrix
    \foreach \x in {1,2,...,\nm}
    {
    \matrix [draw, % for the rectangle border
             fill=white, % so that it is not transparent
             ampersand replacement=\&] %see explanation
    (mm\x)%give the matrix a name
    at(-\x * \xs, -\x * \ys) %shift the matrix
    {
        \node {(\x,1,1)}; \& \node {(\x,2,1)}; \& \node {(\x,3,1)}; \& \node {(\x,4,1)};\\
        \node {(\x,1,2)}; \& \node {(\x,2,2)}; \& \node {(\x,3,2)}; \& \node {(\x,4,2)};\\
    };
    }

    \draw [dashed,gray](mm1.north west) -- (mm\nm.north west);
    \draw [dashed,gray](mm1.north east) -- (mm\nm.north east);
    \draw [dashed,gray](mm1.south east) -- (mm\nm.south east);
    \end{tikzpicture}
\end{figure}

\subsubsection{Linearized Indexing} \label{linearizedindexing}
Note that in $E$, elements are ordered in ascending order with respect to their indexing. First, sub-array $(1)$ is populated, and $(1, 1)$ is the first array within it, so it is filled with elements $1$ and $2$. Array $(1, 2)$ is then populated by elements $3$ and $4$. After all four of \textbf{$E$}'s sub-arrays have been populated, sub-array $(2)$ is populated, and so on until the array is full.

Given the numbering scheme in $E$, one can imagine a ``flattened" version of $E$ that is simply an ordered one-dimensional array containing all of $E$'s elements; simply put, an array containing the elements $1$ to $24$, in numerical order. \matlab\@ actually allows arrays of arbitrarily many dimensions to be logically ``flattened" during accesses so that elements can be accessed with one index, like in a one-dimensional array. We found this property to be somewhat interesting and challenging to model, and we refer to the ``flattening" of a matrix as ``linearization".

The process of linearizing a matrix is perhaps
better explained with a two-dimensional example. Imagine a general two-dimensional $3 \times 4$ matrix, such as the one below, where:
\begin{equation*}
    \begin{bmatrix}
        a & b & c & d \\
        e & f & g & h \\
        i & j & k & l
    \end{bmatrix}
\end{equation*}
To linearize the matrix, we label each value with an increasing index from the
top-left, left-to-right, top-to-bottom. For example:
\begin{equation*}
    \begin{bmatrix}
        a_1 & b_2 & c_3 & d_4 \\
        e_5 & f_6 & g_7 & h_8 \\
        i_9 & j_{10} & k_{11} & l_{12}
    \end{bmatrix}
\end{equation*}
Finally, we create a vector, a one-dimensional matrix, by reading the elements
off in order:
\setcounter{MaxMatrixCols}{20}
\begin{equation*}
    \begin{bmatrix}
        a & b & c & d & e & f & g & h & i & j & k & l
    \end{bmatrix}
\end{equation*}
This process generalizes to higher dimensions. The length of the resulting
vector is always the product of the sizes of the dimensions. To see this, note
that this product counts the number of elements in the matrix and that the
linearized matrix has the same number of elements in a single dimension.

Under this scheme, any element in a matrix can be accessed via its position in the linearized version of that matrix. For example, the element $12$ in \textbf{$E$} can be accessed by the index $(12)$ (note that 12 refers to the index, not the elements) and in the previous example, the element $f$ can be accessed by the index $(6)$.

\subsubsection{Ranged Indexing} \label{rangedindexing}
Broadly speaking, \matlab\@ supports the notion of a range of numbers specified by a starting number and a stopping number, which includes every number between the starting number and the stopping number (inclusive). This abstraction can be used to access or assign elements within a sub-matrix of a particular matrix. For example, suppose in matrix $E$ you only cared about the first matrix in the first dimension, the second/third/fourth columns within the second dimension, and the second row within the third dimension. This sub-matrix can be specified by the index $(1, 2:4, 2)$ and would return the matrix:
\begin{equation*}
    \begin{bmatrix}
        4 & 6 & 8
    \end{bmatrix}
\end{equation*}

Suppose that you wanted to get the second/third/fourth columns within the second dimension and second row within the third dimension for \textit{all three} of the outer-most dimension of \textbf{$E$} but did not know a priori how many elements it had. \matlab\@ supports an ``inclusive range" operator, represented by a colon \textbf{:}, which is taken to mean the range encompassing all values in a particular dimension. The sub-matrix described here would be specified by the index $(:, 2:4, 2)$ and represents the following matrix:

\begin{figure}[h!]
    \centering
    \begin{tikzpicture}[every node/.style={anchor=north east,fill=white, minimum width=.7cm}]
    \matrix (mA) [draw,matrix of math nodes]
    {
    4 & 6 & 8 \\
    };

    \matrix (mB) [draw,matrix of math nodes] at ($(mA.south west)-(-1.5, .1)$)
    {
    12 & 14 & 16 \\
    };

    \matrix (mC) [draw,matrix of math nodes] at ($(mB.south west)-(-1.5, .1)$)
    {
    20 & 22 & 24 \\
    };

    \draw[dashed](mA.north east)--(mC.north east);
    \draw[dashed](mA.north west)--(mC.north west);
    \draw[dashed](mA.south east)--(mC.south east);
    \end{tikzpicture}

\end{figure}

\subsubsection{Matrix Operations} \label{matrixoperations}
Details about \matlab\@ arrays pertaining to indexing are perhaps some of their most interesting features, but we also use some typical matrix operators in our program. Generally speaking, matrix operators can be generalized to any number of dimensions, as long as the dimensions of the arrays agree for the operation in question. There are two broad classes of matrix operators that we make use of:

\begin{itemize}
    \item \textbf{Element-wise}: For two matrices of the same dimensionality, the operator is applied on each of the two elements containing the same index in the two matrices.
    \item \textbf{All Elements}: For an array and a scalar, the operation and scalar (e.g. add 1) is applied to every element in the matrix.
\end{itemize}

\subsection{Modeling the Problem}
Aside from the core matrix operations, we anticipated that we would need to model the state of our program at all points of execution since our goal was to prove termination. To this end, we would need to identify all of the syntax that we use in our program and specify their associated semantics. The upshot of modeling our program in this way would be the ability to remains to prove its termination of our given that certain preconditions hold.

Hoare logic~\cite{Hoare_1969} is a type of logic that is often used to solve problems of this class. Partial Hoare logic is a weak statement about programs; namely, given preconditions \(P\) and postconditions \(Q\) and a program \(C\), the Hoare triple \(\{P\}C\{Q\}\) asserts that \emph{if} \(C\) terminates, then the pre-conditions imply the post-conditions. Total Hoare logic, by contrast, is a stronger version of this claim: \(C\) is asserted to terminate. Thus a proof in total Hoare logic necessitates a proof that \(C\) terminates; the inference rules are accordingly adjusted. The primary adjustment is to the rule for while
loops, which are guaranteed to terminate by relation to a decreasing sequence
(with respect to a well-founded order).

When we started this project, we had in mind that the preconditions would be statements about the program's inputs; for example, ``the length and width of \textit{key art} are always multiples of the \textit{thumbnails'} size". Ultimately, we did not get far enough in modeling our program to devise a set of pre-conditions, but the point remains that we needed to be able to model the control flow of our entire program in order to do this.

\section{Design / Approach}

We first set out to define the syntax and semantics of the subset of \matlab\ used in our program, making the assumption that once we finished this it would be somewhat trivial to bind the defined syntax to behavior in Coq. Ultimately, the majority of our effort in this project ended going towards the implementation of our multi-dimensional matrix operations, which was made difficult by the number of properties that our matrices uphold as well as our desire to write generic Coq functions that would work on matrices of arbitrarily many dimensions. We spent a while working with Coq's built-in \textsf{Vector} type before coming to the conclusion that its dependent typing caused difficulties in behavior specification. To this end, we ended up implementing our own data types to describe the state of matrices in our program; this task was unexpectedly challenging and once we began working on it, it became the only thing we worked on for the remainder of the project.

@@

We present a brief overview on \mmatlab, the subset of \matlab\ that we modeled in
order to prove facts about our programs.

Aside from general programming patterns for control flow, the main syntax of
relevance in our program involves array/matrix accesses and updates, arithmetic
operations, and \matlab-specific primitives such as ranges. We have identified
all of the expressions and statements we expected to use and defined them as
inductive types, shown in \figurename~\ref{fig:inductivetypeexp} and
\figurename~\ref{fig:inductivetypestmt}.

@@

    \item \textsf{AddExp}: While this is nominally a regular addition operator,
        \matlab's addition is interesting in that it has overloaded parameter
        types and expected behavior for each. The addition of two numbers is a
        number and the addition of two matrices results in element-wise
        addition, as one might expect. However, the sum of a number and a
        matrix, is a matrix with all elements incremented by the number. Similar
        behaviors hold for \textsf{MultExp} and \textsf{DivExp}. See Section \ref{matrixoperations} for the distinction between these operators.

    \item \textsf{RangeExp}: Like Python, \matlab\ has an expression where a
        starting number and stopping number may be provided, generating an
        inclusive list of all numbers between them. This is discussed in Section \ref{rangedindexing}

    \item \textsf{IndexExp}: In support of the RangeExp expression, the syntax to access elements from an array is overloaded in \matlab\@. They

@@

@@

\section{Old / Superceded}

\begin{figure}[!t]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
        \FOR{\(tile \in tiles\)}
            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
            \ENDFOR
            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
            \STATE{\(best\_index \gets best\_indices_0\)}
            \STATE{\(tile \gets thumbnails_{best\_index}\)}
        \ENDFOR
    \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 1}\label{alg:v1}
\end{figure}

\begin{figure}[!t]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
        \STATE{\(used \gets \emptyset\)}
        \STATE{\(threshold \gets 1\)}
        \IF{\(|thumbnails| < |tiles|\)}
            \STATE{\(threshold \gets \lceil{|tiles| / |thumbnails|}\rceil\)}
        \ENDIF
        \FOR{\(tile \in tiles\)}
            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
            \ENDFOR
            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
            \STATE{\(best\_index \gets best\_indices_0\)}
            \WHILE{\(used_{best\_index} \ge threshold\)}
                \STATE{\(mses_{best\_index} \gets \infty\)}
                \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
                \STATE{\(best\_index \gets best\_indices_0\)}
            \ENDWHILE
            \STATE{increment \(used_{best\_index}\)}
            \STATE{\(tile \gets thumbnails_{best\_index}\)}
        \ENDFOR
        \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 2}\label{alg:v2}
\end{figure}

\begin{figure}[!t]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{}\COMMENT{Repeat Version 2}
        \STATE{\(tiles \gets \textrm{Version2}(key\_art, thumbnails, size)\)}
        \STATE{\(tiles \gets \textrm{imgaussfilt}(tiles)\)}
        \STATE{\(strip\_size \gets \lfloor{size / 8}\rfloor\)}
        \STATE{}\COMMENT{Blur by averaging}
        \FOR{horizontal \(strip \in tiles\)}
            \STATE{\(strip \gets \textrm{mean}({strip})\)}
        \ENDFOR
        \FOR{vertical \(strip \in tiles\)}
            \STATE{\(strip \gets \textrm{mean}({strip})\)}
        \ENDFOR
        \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 3}\label{alg:v3}
\end{figure}

\section{Goals and Timeline}

As we neglected to include them in our original proposal, we now propose the
following goals, taking into consideration the progress that we have made so far
with the project:

\begin{enumerate}

    \item \textbf{Minimum Viable Product}: Successful modelling of a majority of
        the desired features corresponding to matrix operations in \mmatlab\@.
        Given the lack of multi-dimensional matrix libraries in Coq and the
        necessary additions of several features in the Coq standard library,
        this alone would be a notable achievement for us.

    \item \textbf{Desired Goal}: Successful binding of our matrix operations to
        \mmatlab\ statements and expressions, including the ability to evaluate
        programs written in \mmatlab\@. A proof of termination, or sketch of
        one, for our V1 algorithm is provided.

    \item \textbf{Stretch Goal}: All of the above, and the addition of all
        operations used in V2 and V3 algorithms (most notably, for loops that
        increment across a range at varying intervals). Proofs that use
        these features.

\end{enumerate}

Given what we have encountered so far, we propose the following modified
timeline for project completion.

\begin{itemize}
    \item \textbf{03/31/21}: Progress Report Due
    \item \textbf{04/09/21}: Significant progress made on most matrix operations
        and data structures.
    \item \textbf{04/16/21}: Binding \mmatlab\ syntax to operations written
        in Coq; definition of semantics and program evaluator; formal proof
        specification written.
    \item \textbf{04/24/21}: Proof(s) should be written or scope adjusted
        accordingly. All logic should be implemented; edits for readability
        should be the main changes henceforth.
    \item \textbf{04/26/21}: Final Proof Due
    \item \textbf{05/03/21}: Final Paper Due
    \item \textbf{05/05/21}: Final Presentation Delivered
\end{itemize}

We provide the following ``diff'' between the scope of our original project
proposal and our new proposal:
\begin{itemize}
    \item Heightened emphasis on implementation of matrix operations, since they
        are proving to be non-trivial.
    \item Addition of goal tiers.
    \item Discussion on Hoare logic.
\end{itemize}
